
\documentclass[11pt]{article} 
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amssymb} 
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=red]{hyperref}
\usepackage{url}

%\usepackage{quiver}


%\setlist[itemize]{label=(\roman*)}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{example}
\newtheorem{example}{Example}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}[section]

\theoremstyle{proposition}
\newtheorem{proposition}{Proposition}[section]

\theoremstyle{Problem}
\newtheorem{problem}{Problem}[section]

\theoremstyle{Solution}
\newtheorem{solution}{Solution}[section]

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}[section]

\theoremstyle{corollary}
\newtheorem{corollary}{Corollary}[section]


\title{Problem Set 1}
\author{Following Samy's talk}
\date{}
\begin{document}

\maketitle
\begin{problem}
Let $(X,\mathcal{M},\mu)$ be a measure space. Given $A,B$ be measurable sets, show that $$\mu(A\cup B) = \mu(A)+\mu(B) - \mu(A\cap B)$$
\end{problem}

\begin{problem}
Let $(X,\mathcal{M},\mu)$ be a measure space. Show that $\mu$ is \emph{monotone}:
$$A\subset B \implies \mu(A) \leq \mu(B)$$
\end{problem}
This problem shows that if $\mu$ is finite (i.e. $\mu(X) <\infty$), every set has finite measure.

\begin{problem}
Show that $\mu$ is continuous. That is, if $A_1\subset ... \subset A_n \subset ...$ is a countable chain, then $$\mu\left(\bigcup_{n=1}^{\infty} A_n\right) = \lim_{n\to \infty} \mu(A_n)$$
Similarily, if $A_1 \supset  ... \supset A_n \supset ...$ is a countable chain and $\mu(A_1)<\infty$, then
$$\mu\left(\bigcap_{n=1}^{\infty} A_n\right) = \lim_{n\to \infty} \mu(A_n)$$
\end{problem}

\begin{problem}
Show that the following sets have Lebesgue measure $0$:

(i) $\mathbb{Q}$ 

(ii) The Cantor set. (Hint: Continuity)

\end{problem}

\begin{problem}
Let $F$ be a finite set. Verify that the counting measure 
$$\mu(A) = |A|$$
is indeed a measure. ($|A|$ denotes the cardinality of the set).
\end{problem}


\end{document}